---
title: "Ordinal Logistic Regression"
author: "Lian"
date: "2016年8月18日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
Below are the steps to perform OLR in R:
[Reference HTTP](https://www.analyticsvidhya.com/blog/2016/02/multinomial-ordinal-logistic-regression/)

```{r Load the Libraries}
require(foreign)
require(ggplot2)
require(MASS)
require(Hmisc)
require(reshape2)
```

```{r Load the data}
dat <- read.dta("http://www.ats.ucla.edu/stat/data/ologit.dta")
head(dat)
```
Let’s quickly understand the data.

The data set has a dependent variable known as apply. It has 3 levels namely “unlikely”, “somewhat likely”, and “very likely”, coded in 1, 2, and 3 respectively. 3 being highest and 1 being lowest. This situation is best for using ordinal regression because of presence of ordered categories. Pared (0/1) refers to at least one parent has a graduate degree; public (0/1) refers to the type of undergraduate institute.

For building this model, we will be using the polr command to estimate an ordered logistic regression. Then, we’ll specify Hess=TRUE to let the model output show the observed information matrix from optimization which is used to get standard errors.

```{r Ordinal Logit model}
m <- polr(apply ~ pared + public + gpa, data = dat, Hess=TRUE)
summary(m)
```

We see the usual regression output coefficient table including the value of each coefficient, standard errors, t values, estimates for the two intercepts, residual deviance and AIC. AIC is the information criteria. Lesser the better.

Now we’ll calculate some essential metrics such as p-Value, CI, Odds ratio

```{r calculate some essential metrics}
ctable <- coef(summary(m))
ctable
p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
ctable <- cbind(ctable, "p value" = p)
ctable
ci <- confint(m)
ci  # confidence intervals
exp(coef(m))
exp(cbind(OR = coef(m), ci))
```
### Interpretation
1. One unit increase in parental education, from 0 (Low) to 1 (High), the odds of “very likely” applying versus “somewhat likely” or “unlikely” applying combined are 2.85 greater .

2. The odds “very likely” or “somewhat likely” applying versus “unlikely” applying is 2.85 times greater .

3. For gpa, when a student’s gpa moves 1 unit, the odds of moving from “unlikely” applying to “somewhat likely” or “very likley” applying (or from the lower and middle categories to the high category) are multiplied by 1.85.

Let’s now try to enhance this model to obtain better prediction estimates.

```{r}
summary(m)
summary(update(m, method = "probit", Hess = TRUE), digits = 3)
summary(update(m, method = "logistic", Hess = TRUE), digits = 3)
summary(update(m, method = "cloglog", Hess = TRUE), digits = 3)
```
Let’s add interaction terms here
```{r}
head(predict(m, dat, type = "p"))
addterm(m, ~.^2, test = "Chisq")
m2 <- stepAIC(m, ~.^2)
m2
summary(m2)
m2$anova
anova(m, m2)
```

```{r to plot this model}
m3 <- update(m, Hess=TRUE)
pr <- profile(m3)
confint(pr)
plot(pr)
pairs(pr)
```


