---
title: "R for Data Science"
author: "Lian"
date: "2016年12月14日"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: readable
    highlight: tango
    code_folding: show
    df_print:
      max.print: 10
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

[R for Data Science](http://r4ds.had.co.nz/)

## 1. Introduction
The majority of the packages that you will learn in this book are part of the so-called tidyverse. The packages in the tidyverse share a common philosophy of data and R programming, and are designed to work together naturally.

```{r message=FALSE, warning=FALSE}
# install.packages("tidyverse")
library(tidyverse)
```

In this book we’ll use three data packages from outside the tidyverse:
```{r}
# install.packages(c("nycflights13", "gapminder", "Lahman"))
```

These packages provide data on airline flights, world development, and baseball that we’ll use to illustrate key data science ideas.

> Explore




