---
title: "Method Comparison"
author: "Lian"
date: "2016年12月6日"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: readable
    highlight: tango
    code_folding: show
    df_print:
      max.print: 10
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## **1. Data: logsouvenirtimeseries**
### 1.1 Simple Moving Average
```{r}
library(TTR)
souvenir <- scan("http://robjhyndman.com/tsdldata/data/fancy.dat")
souvenirtimeseries <- ts(souvenir, frequency=12, start=c(1987,1))
logsouvenirtimeseries <- log(souvenirtimeseries)
plot.ts(logsouvenirtimeseries)
logsouvenirtimeseriesSMA3 <- SMA(logsouvenirtimeseries,n=3)
logsouvenirtimeseriesSMA5 <- SMA(logsouvenirtimeseries,n=5)
ts.plot(logsouvenirtimeseries, logsouvenirtimeseriesSMA3, logsouvenirtimeseriesSMA5, gpars = list(col = c("black", "red", "blue")))
```

### 1.2 Exponenial Smoothing
```{r}
library(forecast)
souvenirtimeseriesforecasts <- HoltWinters(logsouvenirtimeseries)
souvenirtimeseriesforecasts
plot(souvenirtimeseriesforecasts)
souvenirtimeseriesforecasts2 <- forecast.HoltWinters(souvenirtimeseriesforecasts, h=24)
plot.forecast(souvenirtimeseriesforecasts2)
```

### 1.3 ARIMA
```{r}
library(forecast)
arima <-auto.arima(logsouvenirtimeseries,ic='bic',trace=T)
forecast <- forecast.Arima(arima, h=24, level=c(99.5))
plot.forecast(forecast)
```

### 1.4 Wavelet
```{r}
library(WaveletComp)
wt = analyze.wavelet(as.data.frame(logsouvenirtimeseries), my.series=1,loess.span=0.75,dt=1, dj=1/250,lowerPeriod=1,upperPeriod=64,make.pval=T, n.sim=10)

wt.image(wt, color.key = "quantile", n.levels = 100, legend.params = list(lab = "wavelet power levels", mar = 4.7))
reconstruct(wt, plot.waves = F, lwd = c(1,2), show.legend = FALSE, legend.coords = "topleft")
```

## **2. Data: skirtsseries**
### 2.1 Simple Moving Average
```{r}
skirts <- scan("http://robjhyndman.com/tsdldata/roberts/skirts.dat",skip=5)
skirtsseries <- ts(skirts,start=c(1866))
plot.ts(skirtsseries)
skirtsseriesSMA3 <- SMA(skirtsseries,n=3)
skirtsseriesSMA5 <- SMA(skirtsseries,n=5)
ts.plot(skirtsseries, skirtsseriesSMA3, skirtsseriesSMA5, gpars = list(col = c("black", "red", "blue")))
```

### 2.2 Exponenial Smoothing
```{r}
skirtsseriesforecasts <- HoltWinters(skirtsseries, gamma=FALSE)
skirtsseriesforecasts
plot(skirtsseriesforecasts)
skirtsseriesforecasts2 <- forecast.HoltWinters(skirtsseriesforecasts, h=12, gamma=FALSE)
plot.forecast(skirtsseriesforecasts2)
```

### 2.3 ARIMA
```{r}
arima <-auto.arima(skirtsseries,ic='bic',trace=T)
forecast <- forecast.Arima(arima, h=12, level=c(99.5))
plot.forecast(forecast)
```

### 2.4 Wavelet
```{r}
wt = analyze.wavelet(as.data.frame(skirtsseries), my.series=1,loess.span=0, dt=1, dj=1/250,lowerPeriod=1,upperPeriod=256,make.pval=T, n.sim=10)

wt.image(wt, color.key = "quantile", n.levels = 100, legend.params = list(lab = "wavelet power levels", mar = 4.7))
reconstruct(wt, plot.waves = F, lwd = c(1,2), show.legend = FALSE, legend.coords = "topleft")
```

## 3. **Data: volcanodustseries**
### 3.1 Simple Moving Average
```{r}
volcanodust <- scan("http://robjhyndman.com/tsdldata/annual/dvi.dat", skip=1)
volcanodustseries <- ts(volcanodust,start=c(1500))
plot.ts(volcanodustseries)
volcanodustseriesSMA3 <- SMA(volcanodustseries,n=3)
volcanodustseriesSMA5 <- SMA(volcanodustseries,n=5)
ts.plot(volcanodustseries, volcanodustseriesSMA3, volcanodustseriesSMA5, gpars = list(col = c("black", "red", "blue")))
```

### 3.2 Exponenial Smoothing
```{r}
volcanodustseriesforecasts <- HoltWinters(volcanodustseries,beta=FALSE, gamma=FALSE)
volcanodustseriesforecasts
plot(volcanodustseriesforecasts)
volcanodustseriesforecasts2 <- forecast.HoltWinters(volcanodustseriesforecasts, h=12, beta=FALSE, gamma=FALSE)
plot.forecast(volcanodustseriesforecasts2)
```

### 3.3 ARIMA
```{r}
arima <-auto.arima(volcanodustseries,ic='bic',trace=T)
forecast <- forecast.Arima(arima, h=12, level=c(99.5))
plot.forecast(forecast)
```

### 3.4 Wavelet
```{r}
wt = analyze.wavelet(as.data.frame(volcanodustseries), my.series=1,loess.span=0, dt=1, dj=1/250,lowerPeriod=1,upperPeriod=64,make.pval=T, n.sim=10)

wt.image(wt, color.key = "quantile", n.levels = 100, legend.params = list(lab = "wavelet power levels", mar = 4.7))
reconstruct(wt, plot.waves = F, lwd = c(1,2), show.legend = FALSE, legend.coords = "topleft")
```

## **4. cust_no 341003060003518**
```{r}
customer1 <- readRDS('cust_no_341003060003518.rds')
customer1 <- ts(customer1$spending, frequency=1, start=2)
```

### 4.1 Simple Moving Average
```{r}
plot.ts(customer1)
customer1SMA3 <- SMA(customer1,n=3)
customer1SMA5 <- SMA(customer1,n=5)
ts.plot(customer1, customer1SMA3, customer1SMA5, gpars = list(col = c("black", "red", "blue")))
```

### 4.2 Exponenial Smoothing
```{r}
customer1forecasts <- HoltWinters(customer1, gamma=FALSE)
customer1forecasts
plot(customer1forecasts)
customer1forecasts2 <- forecast.HoltWinters(customer1forecasts, h=4, gamma=FALSE)
plot.forecast(customer1forecasts2)
```

### 4.3 ARIMA
```{r}
arima <-auto.arima(customer1,ic='bic',trace=T)
acf(customer1, lag.max = 20)
pacf(customer1, lag.max = 20)
forecast <- forecast.Arima(arima, h=4, level=c(99.5))
plot.forecast(forecast)
```

### 4.4 Wavelet
```{r}
wt = analyze.wavelet(as.data.frame(customer1), my.series=1,loess.span=1, dt=1, dj=1/250,lowerPeriod=1,upperPeriod=64,make.pval=T, n.sim=10)

wt.image(wt, color.key = "quantile", n.levels = 100, legend.params = list(lab = "wavelet power levels", mar = 4.7))
reconstruct(wt, plot.waves = F, lwd = c(1,2), show.legend = FALSE, legend.coords = "topleft")
```

## **5. cust_no 341003060065545**
```{r}
customer2 <- readRDS('cust_no_341003060065545.rds')
customer2 <- ts(customer2$spending, frequency=1, start=2)
```

### 5.1 Simple Moving Average
```{r}
plot.ts(customer2)
customer2SMA3 <- SMA(customer2,n=3)
customer2SMA5 <- SMA(customer2,n=5)
ts.plot(customer2, customer2SMA3, customer2SMA5, gpars = list(col = c("black", "red", "blue")))
```

### 5.2 Exponenial Smoothing
```{r}
customer2forecasts <- HoltWinters(customer2, gamma=FALSE)
customer2forecasts
plot(customer2forecasts)
customer2forecasts2 <- forecast.HoltWinters(customer2forecasts, h=4, gamma=FALSE)
plot.forecast(customer2forecasts2)
```

### 5.3 ARIMA
```{r}
arima <-auto.arima(customer2,ic='bic',trace=T)
acf(customer2, lag.max = 20)
pacf(customer2, lag.max = 20)
forecast <- forecast.Arima(arima, h=4, level=c(99.5))
plot.forecast(forecast)
```

### 5.4 Wavelet
```{r}
wt = analyze.wavelet(as.data.frame(customer2), my.series=1,loess.span=1, dt=1, dj=1/250,lowerPeriod=1,upperPeriod=64,make.pval=T, n.sim=10)

wt.image(wt, color.key = "quantile", n.levels = 100, legend.params = list(lab = "wavelet power levels", mar = 5.7))
reconstruct(wt, plot.waves = F, lwd = c(1,2), show.legend = FALSE, legend.coords = "topleft")
```